Moral Foundations of U.S. Political News Organizations

### 2019-02-09 pilot_manuscript.tex
idk i reinstalled papaja??? also, small update to citations

\authornote{Add complete departmental affiliations for each author here. Each new line herein must be indented, like this line.

William Padfield is a master's degree candidate in psychology at Missouri State University. This thesis partially fulfills the requirements for the Master of Science degree in Psychology.

@@

\abstract{
Enter abstract here. Each new line herein must be indented, like this line.

@@

\hypertarget{moral-foundations-theory}{%
\subsection{Moral Foundations Theory}\label{moral-foundations-theory}}

As a discipline, modern moral psychology started in the late 1960s with Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research popularized his theory of the development of moral reasoning. This theory establishes the steps of moral reasoning through which humans proceed as their cognitive structures assume higher levels of sophistication and nuance ({\textbf{???}}). Kohlberg borrowed from Jean Piaget's stages of cognitive development in which children progress from the sensorimotor through to the formal operations stage. Similarly, Kohlberg found people typically start with a \enquote{pre-conventional} understanding of morality during infancy in which children understand \enquote{right} and \enquote{wrong} purely in terms of how they interact with resultant experiences of rewards and punishment. Typically, people progress through several steps until they reach a \enquote{post-conventional} ethics. People who have reached the post-conventional stage are said to be able to weigh competing abstractions and reason their way to a conclusion that promotes justice based upon their \enquote{self-chosen ethical principles} ({\textbf{???}}). From Kohlberg's perspective, issues of justice and fairness comprise the foundation of morality (Haidt \& Graham, 2007). This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the grounds that it focused solely on the moral concerns of men, and that it ignored those of women (Haidt \& Graham, 2007). Gilligan drew attention to purported differences in the ways men and women are taught to relate to self and others. She offered a historic argument contending women have traditionally filled roles related to caring and nurturing. She pushes back against Kohlberg's assumption that moral development replaces \enquote{rule of brute force,} as enforced by men, with the justice-based \enquote{rule of law.} According to Gilligan, this assumption implies women are less morally developed, owing to their absence both in masculine displays of violence as well as in enforcement of the law ({\textbf{???}}). Gilligan argues for the existence of a distinct, but equal development process that women and girls must undergo in order to develop their moral selves. Stark differences in the ways women are traditionally taught to interact with their social world cause them to develop ethical systems based upon their non-aggressive relationships with others. Gilligan thus asserted morality was built upon an alternative moral foundation: caring ({\textbf{???}}). This debate between competing conceptions of morality did not resolve until Gilligan and Kohlberg conceded the existence of two moral foundations: justice and caring (Haidt \& Graham, 2007). While this new direction in moral psychology appeared to represent a more inclusive outlook on the construct, these novel ideas would soon be challenged on the grounds of its apparent western-centric outlook.

@@

\hypertarget{moral-foundations-dictionary}{%
\subsection{Moral Foundations Dictionary}\label{moral-foundations-dictionary}}

@@

\hypertarget{valence}{%
\subsection{Valence}\label{valence}}

@@

In order to source a large number of lemmas for affective ratings, the researchers drew from several validated sources. These include the 30,000 lemmas with age-of-acquisition (average age at which a particular word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US corpus consisting of subtitles from various forms of visual media ({\textbf{???}}). This data collection resulted in the final list of 13,915 lemmas. Lists of 346-350 words were presented to participants recruited through the Amazon Mechanical Turk subject pool. Participants rated the words along one of the three dimensions, unlike the ANEW project in which participants rated each word along all three dimensions at once. The researchers used a nine-point scale similar to the one used by Bradley and Lang (1999) when collecting ratings for ANEW (Warriner et al., 2013).

@@

Warriner and Kuperman (2015) utilized the new affective norms list in order to investigate the validity of the Pollyanna hypothesis, or the prevalence of a generally optimistic outlook in humans as reflected in language. The researchers were able to conclude the existence of a greater number of positive-valence English words in the list of 13,915 lemmas. Additionally, after observing token frequency in a number of text corpora, including SUBTLEX-US, COCA, BNC, TASA, and HAL, the researchers found that words with positive valence were also used more frequently (Warriner \& Kuperman, 2015). While the researchers concede the possibility of an acquiescence bias in ratings as a possible explanation for the observed positivity bias, this investigation represents one application of the Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner et al. (2013) norms have been utilized in cognitive research as well. One cognition-based study investigates the relationship between emotion and response latencies in word recognition. Kuperman, Estes, Brysbaert, and Warriner (2014) sought to use these new norms to fill in the knowledge gaps regarding variance in word recognition. The researchers drew several conclusions regarding emotion and word recognition (specifically in naming and lexical decision tasks). First, Kuperman et al. (2014) found slower decision-making and reading times in negative-valence words, faster times in neutral words, and even faster times in words with positive valence. The researchers also concluded that words causing higher arousal tend to have slower decision times than less-arousing words. They found valence had a stronger effect on recognition than arousal (both effects were independent, not interactive). They found an interaction between emotion and word frequency such that valence and arousal are more effective on lower frequency words than high frequency words. Finally, Kuperman et al. (2014) found a greater effect of valence and arousal on response latency for lexical decision tasks than for naming tasks (Kuperman et al., 2014). This research serves as further evidence that the Warriner et al. (2013) list can be used for research inquiries both within and without the field of psycholinguistics.

@@

\hypertarget{news-media-and-politics}{%
\subsection{News Media and Politics}\label{news-media-and-politics}}

@@

They linked the Nixon administration's awareness of the issues for which the president had public support to the issues he should emphasize (and prime) during the campaign. Likewise the researchers found evidence that Nixon's team was aware of negative evaluations of his warmth and trustworthiness, and thus took steps to prime his purportedly positive qualities, including strength and competence (Druckman et al., 2004). The researchers also cited research from Iyengar and Kinder (1987) suggesting the news media affected perceptions of President Jimmy Carter's competence by emphasizing (e.g.~priming) issues related to energy, defense, and the economy. This focus implies news media may contribute to Americans' perception of politicians based on where the media places emphasis.

There is a potential caveat regarding the validity of Druckman et al. (2004)'s findings: reproductions of several studies purporting to demonstrate social priming effects have failed to replicate the original results. ({\textbf{???}}) point out the distinction between perceptual and social (or goal) priming both in their operational definitions as well as their replicability. Perceptual priming often works through the inducement of a certain response from a related prime, as in, for example, semantic priming. Social (or goal) priming encompasses phenomena by which people exhibit complex behavioral changes subsequent to exposure to a prime. ({\textbf{???}}) point out well-known studies investigating social priming, including the use of elderly-related primes to induce slower walking speeds in participants. Studies investigating perceptual priming have been \enquote{directly replicated in hundreds of labs} ({\textbf{???}}). This does not appear to be the case for social priming, as argued by ({\textbf{???}}).

({\textbf{???}}) noticed the unusually large effect size values (Cohen's \emph{d}) reported by researchers studying social priming effects. The researchers reproduced two studies from ({\textbf{???}}) The first study attempted to prime participants by having them plot points on a Cartesian grid. The independent variable was priming condition and contained three levels: short, middle, and long distance. Those instructed to plot points further apart were hypothesized to express a higher degree of psychological distance regarding their family. The second study used the same priming conditions, but hypothesized that greater distance between points would prime participants to estimate fewer calories in unhealthy foods than those who were primed with shorter distances between points. ({\textbf{???}}) concluded those two studies from ({\textbf{???}}) held little validity while also casting doubt on the prevalence of social priming effects themselves, based on the inability of other researchers to replicate previously reported effects in this area.

@@

\hypertarget{experiment-1}{%
\section{Experiment 1}\label{experiment-1}}

\hypertarget{method}{%
\section{Method}\label{method}}

@@

\hypertarget{sources}{%
\subsection{Sources}\label{sources}}

@@

\hypertarget{materials}{%
\subsection{Materials}\label{materials}}

Using the \emph{rvest} library in the statistical package \emph{R}, we pulled body text for individual articles from each of the aforementioned sources (identified using CSS language) and compiled them into a dataset ({\textbf{???}}). Using this dataset, we identified word count and average word count per source. This process was run once daily starting on \emph{DATE} until \emph{DATE.} Starting on \emph{DATE,} the process was run twice daily - once in the morning and again in the evening. Data collection was terminated once 250,000 words per source was collected on \emph{DATE.}

\hypertarget{data-analysis}{%
\subsection{Data analysis}\label{data-analysis}}

@@

Concurrent research by Jordan, Buchanan \& Padfield is assessing the validity of both the Moral Foundations Questionnaire and the Moral Foundations Dictionary through a multi-trait multi-method analysis of the two instruments using one sample. The instruments will be analyzed against one another, in order to test reliability, as well as against the Congressional Record in order to test predictive validity for political orientation. The researchers were able to identify a number of potential new words that, if added to the MFD, could comprise a dictionary with greater validity. Additionally, the researchers are investigating whether or not a reduced set of MFD words may likewise bolster the validity of the MFD. Depending on the results of that research, either an augmented or reduced Moral Foundations Dictionary may be utilized in this analysis rather than the original version.

@@

\hypertarget{results}{%
\section{Results}\label{results}}

@@

The multilevel model did not indicate the presence of any significant or practical effect of political lean for any of the five moral foundations. The strongest effect size was observed for the \emph{authority/respect} foundation, but the effect was in the opposite direction from what was originally hypothesized - liberal sources tended to use more \emph{authority/respect} words than did conservative sources. Descriptive and test statistics, \emph{p}-values and effect sizes (Cohen's \emph{d}) can be found in Table 1. Based on the weighted percent values for the five foundations, the researchers observed that MFD words seem to make up a small portion of the article text. Furthermore, the observed percentages and means appear to indicate a generally positive endorsement of all five foundations across both liberal and conservative sources.

@@

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\hypertarget{experiment-2}{%
\section{Experiment 2}\label{experiment-2}}

\hypertarget{kavanaugh-supreme-court-hearing}{%
\subsection{Kavanaugh Supreme Court Hearing}\label{kavanaugh-supreme-court-hearing}}

In the wake of Justice Anthony Kennedy's retirement from the Supreme Court of the United States, President Donald Trump nominated Brett Kavanaugh as the new Associate Justice. Kavanaugh was previously on the U.S. Court of Appeals for the District of Columbia. The Senate Judiciary Committee began his confirmation hearing on September 4, 2018 ({\textbf{???}}). Following allegations of sexual assault by high school classmate Dr.~Christine Blasey Ford, the committee postponed its vote on whether or not to open the confirmation to the entire Senate.

On September 27, the committee questioned Dr.~Ford before commencing a second round of questioning for Judge Kavanaugh ({\textbf{???}}). During the intervening weeks between hearings, two more women came forward with two separate allegations of sexual assault on the part of Kavanaugh. According to Nielsen reports, more than 20 million people watched the September 27 proceedings on television (O'Connell, 2018). This figure does not take into account viewers who watched online, nor does it account for viewers outside the United States. On September 28, the Senate Judiciary Committee voted to send the nomination to the Senate floor. Senator Jeff Flake of Arizona, however, lobbied for a week-long FBI investigation on Kavanaugh and the allegations facing him, which the committee, and later the President, approved. The investigation concluded with no significant findings. The Senate voted 50-48 to approve Kavanaugh's appointment on October 6, 2018 ({\textbf{???}}).

@@

\hypertarget{method-1}{%
\section{Method}\label{method-1}}

\hypertarget{sources-1}{%
\subsection{Sources}\label{sources-1}}

@@

\hypertarget{materials-1}{%
\subsection{Materials}\label{materials-1}}

Using the \emph{rvest} library in the statistical package \emph{R}, we pulled body text for individual articles from each of the aforementioned 10 news sources (identified using CSS language). We compiled the articles into a dataset ({\textbf{???}}). Using this dataset, we identified word count and average word count per source. This process was run for articles pertaining to Kavanaugh's nomination that were published between September 13, 2018 and October 11, 2018 inclusive. This date range was selected in reference to the widely-publicized and viewed nomination hearing on September 27, 2018. We set the start date at September 13 (two weeks before the hearing) and the end date at October 11 (two weeks after the hearing) so that we could capture a large amount of data (roughly one month) during which Kavanaugh's nomination was at its peak saturation in news coverage.

\hypertarget{data-analysis-1}{%
\subsection{Data analysis}\label{data-analysis-1}}

As in Experiment 1, the text was scanned with \emph{ngram}. Again, blank articles, text from the Disqus system, and duplicate articles were removed (Schmidt et al., 2017). The text was processed and stemmed in order to convert to a usable form for further analysis (Feinerer \& Hornik, 2017). Words were subsequently matched with their valences from Warriner et al. (2013). Depending on the results of Jordan, Buchanan \& Padfield's research, alternative forms of the Moral Foudnations Dictionary may be imported instead of the original dictionary.

@@

\hypertarget{experiment-2-results}{%
\section{Experiment 2 results}\label{experiment-2-results}}

\hypertarget{discussion-1}{%
\section{Discussion}\label{discussion-1}}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

How it turned out

Limitations of this one (no rationale behind choosing sources except \enquote{people say NPR's liberal??})

What to do for future project (focus on one MF? Different sources? More sources?)

Argue: why is still a good thing to study?? (probably something about current state of discourse, information, truth, \enquote{alternative facts,} subjective reality - philosophical stuff)

@@

\hypertarget{references}{%
\section{References}\label{references}}

@@

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Aust2017}{}%
Aust, F., \& Barth, M. (2017). papaja: Create APA manuscripts with R Markdown. Retrieved from \url{https://github.com/crsh/papaja}

\leavevmode\hypertarget{ref-Bradley1999}{}%
Bradley, M. M., \& Lang, P. J. (1999). \emph{Affective Norms for English Words (ANEW): Instruction Manual and Affective Ratings} (No. C-1). The Center for Research in Psychophysiology, University of Florida.

\leavevmode\hypertarget{ref-Cameron2006}{}%
Cameron, D. (2006). Ideology and language. \emph{Journal of Political Ideologies}, \emph{11}(2), 141--152. doi:\href{https://doi.org/10.1080/13569310600687916}{10.1080/13569310600687916}

\leavevmode\hypertarget{ref-Clinton2014}{}%
Clinton, J. D., \& Enamorado, T. (2014). The National News Media's Effect on Congress: How Fox News Affected Elites in Congress. \emph{The Journal of Politics}, \emph{76}(4), 928--943. doi:\href{https://doi.org/10.1017/S0022381614000425}{10.1017/S0022381614000425}

\leavevmode\hypertarget{ref-Druckman2004}{}%
Druckman, J. N., Jacobs, L. R., \& Ostermeier, E. (2004). Candidate Strategies to Prime Issues and Image. \emph{The Journal of Politics}, \emph{66}(4), 1180--1202. doi:\href{https://doi.org/10.1111/j.0022-3816.2004.00295.x}{10.1111/j.0022-3816.2004.00295.x}

\leavevmode\hypertarget{ref-Feinerer2017}{}%
Feinerer, I., \& Hornik, K. (2017). Text mining package. Retrieved from \url{http://tm.r-forge.r-project.org/}

\leavevmode\hypertarget{ref-Gelman2006}{}%
Gelman, A. (2006). Multilevel (hierarchical) modeling: What it can and cannot do. \emph{Technometrics}, \emph{48}(3), 432--435. doi:\href{https://doi.org/10.1198/004017005000000661}{10.1198/004017005000000661}

\leavevmode\hypertarget{ref-Graham2009}{}%
Graham, J., Haidt, J., \& Nosek, B. A. (2009). Liberals and conservatives rely on different sets of moral foundations. \emph{Journal of Personality and Social Psychology}, \emph{96}(5), 1029--1046. doi:\href{https://doi.org/10.1037/a0015141}{10.1037/a0015141}

\leavevmode\hypertarget{ref-Graham2011}{}%
Graham, J., Nosek, B. A., Haidt, J., Iyer, R., Koleva, S., \& Ditto, P. H. (2011). Mapping the moral domain. \emph{Journal of Personality and Social Psychology}, \emph{101}(2), 366--385. doi:\href{https://doi.org/10.1037/a0021847}{10.1037/a0021847}

\leavevmode\hypertarget{ref-Haidt2007}{}%
Haidt, J., \& Graham, J. (2007). When Morality Opposes Justice: Conservatives Have Moral Intuitions that Liberals may not Recognize. \emph{Social Justice Research}, \emph{20}(1), 98--116. doi:\href{https://doi.org/10.1007/s11211-007-0034-z}{10.1007/s11211-007-0034-z}

\leavevmode\hypertarget{ref-Kuperman2014}{}%
Kuperman, V., Estes, Z., Brysbaert, M., \& Warriner, A. B. (2014). Emotion and language: Valence and arousal affect word recognition. \emph{Journal of Experimental Psychology: General}, \emph{143}(3), 1065--1081. doi:\href{https://doi.org/10.1037/a0035669}{10.1037/a0035669}

>>>>

\authornote{ William Padfield is a master's degree candidate
in psychology at Missouri State University. This thesis partially
fulfills the requirements for the Master of Science degree in
Psychology.

@@

\abstract{
The media ecosystem has grown, and political opinions have diverged such
that there are competing conceptions of objective truth. Commentators
often point to political biases in news coverage as a catalyst for this
political divide. The Moral Foundations Dictionary (MFD) facilitates
identification of ideological leanings in text through frequency of the
occurrence of certain words. Through web scraping, the researchers
extracted articles from popular news sources' websites, calculated MFD
word frequencies, and identified words' respective valences. This
process attempts to uncover news outlets' positive or negative
endorsements of certain moral dimensions concomitant with a particular
ideology. In Experiment 1, the researchers gathered poltical articles
from four sources. They were unable to reveal significant differences in
moral or political endorsements, but they solidified the method to be
employed in further research. In Experiment 2, the researchers will
expand their number of sources to 10 and will analyze articles that
pertain to the 2018 confirmation hearings of U.S. Supreme Court Justice
Brett Kavanaugh. This topic was selected due to the moral disputes
associated with his nomination.

@@

\subsection{Moral Foundations Theory}\label{moral-foundations-theory}

As a discipline, modern moral psychology started in the late 1960s with
Lawrence Kohlberg (Haidt \& Graham, 2007). Kohlberg's research
popularized his theory of the development of moral reasoning. This
theory establishes the steps of moral reasoning through which humans
proceed as their cognitive structures assume higher levels of
sophistication and nuance (Kohlberg \& Hersh, 1977). Kohlberg borrowed
from Jean Piaget's stages of cognitive development in which children
progress from the sensorimotor through to the formal operations stage.
Similarly, Kohlberg found people typically start with a
\enquote{pre-conventional} understanding of morality during infancy in
which children understand \enquote{right} and \enquote{wrong} purely in
terms of how they interact with resultant experiences of rewards and
punishment. Typically, people progress through several steps until they
reach a \enquote{post-conventional} ethics. People who have reached the
post-conventional stage are said to be able to weigh competing
abstractions and reason their way to a conclusion that promotes justice
based upon their \enquote{self-chosen ethical principles} (Kohlberg \&
Hersh, 1977). From Kohlberg's perspective, issues of justice and
fairness comprise the foundation of morality (Haidt \& Graham, 2007).
This view persisted until it encountered criticism in the early 1980s.

Kohlberg's conception of morality faced major scrutiny from psychologist
Carol Gilligan. In 1982, Gilligan criticized Kohlberg's theory on the
grounds that it focused solely on the moral concerns of men, and that it
ignored those of women (Haidt \& Graham, 2007). Gilligan drew attention
to purported differences in the ways men and women are taught to relate
to self and others. She offered a historic argument contending women
have traditionally filled roles related to caring and nurturing. She
pushes back against Kohlberg's assumption that moral development
replaces \enquote{rule of brute force,} as enforced by men, with the
justice-based \enquote{rule of law.} According to Gilligan, this
assumption implies women are less morally developed, owing to their
absence both in masculine displays of violence as well as in enforcement
of the law (Gilligan, 1982). Gilligan argues for the existence of a
distinct, but equal development process that women and girls must
undergo in order to develop their moral selves. Stark differences in the
ways women are traditionally taught to interact with their social world
cause them to develop ethical systems based upon their non-aggressive
relationships with others. Gilligan thus asserted morality was built
upon an alternative moral foundation: caring (Gilligan, 1982). This
debate between competing conceptions of morality did not resolve until
Gilligan and Kohlberg conceded the existence of two moral foundations:
justice and caring (Haidt \& Graham, 2007). While this new direction in
moral psychology appeared to represent a more inclusive outlook on the
construct, these novel ideas would soon be challenged on the grounds of
its apparent western-centric outlook.

@@

\subsection{Moral Foundations
Dictionary}\label{moral-foundations-dictionary}

@@

\subsection{Valence}\label{valence}

@@

In order to source a large number of lemmas for affective ratings, the
researchers drew from several validated sources. These include the
30,000 lemmas with age-of-acquisition (average age at which a particular
word is learned) ratings gathered by Kuperman, Stadthagen-Gonzalez, and
Brysbaert (2012) as well as the content lemmas from the SUBTLEX-US
corpus consisting of subtitles from various forms of visual media (New,
Brysbaert, Veronis, \& Pallier, 2007). This data collection resulted in
the final list of 13,915 lemmas. Lists of 346-350 words were presented
to participants recruited through the Amazon Mechanical Turk subject
pool. Participants rated the words along one of the three dimensions,
unlike the ANEW project in which participants rated each word along all
three dimensions at once. The researchers used a nine-point scale
similar to the one used by Bradley and Lang (1999) when collecting
ratings for ANEW (Warriner et al., 2013).

@@

Warriner and Kuperman (2015) utilized the new affective norms list in
order to investigate the validity of the Pollyanna hypothesis, or the
prevalence of a generally optimistic outlook in humans as reflected in
language. The researchers were able to conclude the existence of a
greater number of positive-valence English words in the list of 13,915
lemmas. Additionally, after observing token frequency in a number of
text corpora, including SUBTLEX-US, the Corpus for Contemporary American
English (COCA), the British National Corpus (BNC), Touchstone Applied
Science Associates, Inc. Corpus (TASA), and the corpus used for the
Hyperspace Analogue to Language model (HAL), the researchers found that
words with positive valence were also used more frequently (Warriner \&
Kuperman, 2015). While the researchers concede the possibility of an
acquiescence bias in ratings as a possible explanation for the observed
positivity bias, this investigation represents one application of the
Warriner et al. (2013) list in emotional studies.

In addition to applications in emotion research, the Warriner et al.
(2013) norms have been utilized in cognitive research as well. One
cognition-based study investigates the relationship between emotion and
response latencies in word recognition. Kuperman, Estes, Brysbaert, and
Warriner (2014) sought to use these new norms to fill in the knowledge
gaps regarding variance in word recognition. The researchers drew
several conclusions regarding emotion and word recognition (specifically
in naming and lexical decision tasks - two cognitive processing tasks
wherein a participant has to read aloud or judge a word for its
lexicality). First, Kuperman et al. (2014) found slower decision-making
and reading times in negative-valence words, faster times in neutral
words, and even faster times in words with positive valence. The
researchers also concluded that words causing higher arousal tend to
have slower decision times than less-arousing words. They found valence
had a stronger effect on recognition than arousal (both effects were
independent, not interactive). They found an interaction between emotion
and word frequency such that valence and arousal are more effective on
lower frequency words than high frequency words. Finally, Kuperman et
al. (2014) found a greater effect of valence and arousal on response
latency for lexical decision tasks than for naming tasks (Kuperman et
al., 2014). This research serves as further evidence that the Warriner
et al. (2013) list can be used for research inquiries both within and
without the field of psycholinguistics.

@@

\subsection{News Media and Politics}\label{news-media-and-politics}

@@

They linked the Nixon administration's awareness of the issues for which
the president had public support to the issues he should emphasize (and
prime) during the campaign. Likewise the researchers found evidence that
Nixon's team was aware of negative evaluations of his warmth and
trustworthiness, and thus took steps to prime his purportedly positive
qualities, including strength and competence (Druckman et al., 2004).
The researchers also cited research from Iyengar and Kinder (1987)
suggesting the news media affected perceptions of President Jimmy
Carter's competence by emphasizing (e.g., priming) issues related to
energy, defense, and the economy. This focus implies news media may
contribute to Americans' perception of politicians based on where the
media places emphasis.

There is a potential caveat regarding the validity of Druckman et al.
(2004)'s findings: reproductions of several studies purporting to
demonstrate social priming effects have failed to replicate the original
results. Pashler, Coburn, and Harris (2012) point out the distinction
between perceptual and social (or goal) priming both in their
operational definitions as well as their replicability. Perceptual
priming often works through the inducement of a certain response from a
related prime, as in, for example, semantic priming. Social (or goal)
priming encompasses phenomena by which people exhibit complex behavioral
changes subsequent to exposure to a prime. Pashler et al. (2012) point
out well-known studies investigating social priming, including the use
of elderly-related primes to induce slower walking speeds in
participants. Studies investigating perceptual priming have been
\enquote{directly replicated in hundreds of labs} (Pashler et al.,
2012). This replication rate does not appear to be the case for social
priming, as argued by Pashler et al. (2012).

Pashler et al. (2012) noticed the unusually large effect size values
(Cohen's \emph{d}) reported by researchers studying social priming
effects. The researchers reproduced two studies from Williams and Bargh
(2008) The first study attempted to prime participants by having them
plot points on a Cartesian grid. The independent variable was priming
condition and contained three levels: short, middle, and long distance.
Those instructed to plot points further apart were hypothesized to
express a higher degree of psychological distance regarding their
family. The second study used the same priming conditions, but
hypothesized that greater distance between points would prime
participants to estimate fewer calories in unhealthy foods than those
who were primed with shorter distances between points. Pashler et al.
(2012) concluded those two studies from Williams and Bargh (2008) held
little validity while also casting doubt on the prevalence of social
priming effects themselves, based on the inability of other researchers
to replicate previously reported effects in this area.

@@

Therefore, the current study sought to combine both methods related
questions and extension/replication of previous moral foundation results
found for liberal and conservative sources. First, the MFD was combined
with previous research by the current authors (see below) and weighted
by valence to create weighted percentages to better specify endorsement.
Second, these weighted percentages were examined for their differences
in across liberal and conservative news sources.

\section{Experiment 1}\label{experiment-1}

\section{Method}\label{method}

@@

\subsection{Sources}\label{sources}

@@

\subsection{Materials}\label{materials}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
sources (identified using CSS language) and compiled them into a dataset
(Wickham, 2016). Using this dataset, we identified word count and
average word count per source. This process was run once daily starting
in February 2018 until March 2018. Starting in mid-March 2018, the
process was run twice daily - once in the morning and again in the
evening. Data collection was terminated once 250,000 words per source
was collected in April 2018.

\subsection{Data analysis}\label{data-analysis}

@@

Concurrent research by Jordan, Buchanan, and Padfield (2019) is
assessing the validity of both the Moral Foundations Questionnaire and
the Moral Foundations Dictionary through a multi-trait multi-method
analysis of the two instruments using multiple samples. The instruments
and foundation areas are being analyzed against one another, in order to
test reliability, as well as against the Congressional Record in order
to test predictive validity for political orientation. The researchers
were able to identify a number of potential new words that, if added to
the MFD, could comprise a dictionary with greater validity, and less
likelihood of zero percent texts, as this often occurs with the current
MFD. Those results have informed this analysis, and their updated
findings may change the underlying dictionary used in this analysis
(albeit, we do not expect any changes in the results presented below).

@@

\section{Results}\label{results}

@@

The multilevel model did not indicate the presence of any significant or
practical effect of political lean for any of the five moral
foundations. The strongest effect size was observed for the
\emph{authority/respect} foundation, but the effect was in the opposite
direction from what was originally hypothesized - liberal sources tended
to use more \emph{authority/respect} words than did conservative
sources. Descriptive and test statistics, \emph{p}-values and effect
sizes (Cohen's \emph{d}) can be found in Table 1. To interpret the
weighted scores, one can examine the mean and standard deviations for
each. A zero score for the mean, with a non-zero standard deviation,
would indicate a perfect balance of positive and negative words in each
category, likely representing a neutral tone when all words are
considered. Negative percentages would indicate more representation of
the negative words in the MFD area, while positive percentages indicate
an endorsement of the positive words in a MFD. Therefore, we suggest
using the sign of the mean score to determine the directionality of the
endorsement for the MFD (positive, neutral, negative), and the standard
deviation to ensure that a zero score is not zero endorsement (i.e., a
SD of zero indicates no words were used). Based on the weighted percent
values for the five foundations, the researchers observed that MFD words
seem to make up a small portion of the article text. Furthermore, the
observed percentages and means appear to indicate a generally positive
endorsement of all five foundations across both liberal and conservative
sources.

@@

\section{Discussion}\label{discussion}

@@

\section{Experiment 2}\label{experiment-2}

\subsection{Kavanaugh Supreme Court
Hearing}\label{kavanaugh-supreme-court-hearing}

In the wake of Justice Anthony Kennedy's retirement from the Supreme
Court of the United States, President Donald Trump nominated Brett
Kavanaugh as the new Associate Justice. Kavanaugh was previously on the
U.S. Court of Appeals for the District of Columbia. The Senate Judiciary
Committee began his confirmation hearing on September 4, 2018 (US
Government, 2018a). Following allegations of sexual assault by high
school classmate Dr.~Christine Blasey Ford, the committee postponed its
vote on whether or not to open the confirmation to the entire Senate.

On September 27, the committee questioned Dr.~Ford before commencing a
second round of questioning for Judge Kavanaugh (US Government, 2018b).
During the intervening weeks between hearings, two more women came
forward with two separate allegations of sexual assault on the part of
Kavanaugh. According to Nielsen reports, more than 20 million people
watched the September 27 proceedings on television (O'Connell, 2018).
This figure does not take into account viewers who watched online, nor
does it account for viewers outside the United States. On September 28,
the Senate Judiciary Committee voted to send the nomination to the
Senate floor. Senator Jeff Flake of Arizona, however, lobbied for a
week-long FBI investigation on Kavanaugh and the allegations facing him,
which the committee, and later the President, approved. The
investigation concluded with no significant findings. The Senate voted
50-48 to approve Kavanaugh's appointment on October 6, 2018 (US
Government, 2018c).

@@

\section{Method}\label{method-1}

\subsection{Sources}\label{sources-1}

@@

\subsection{Materials}\label{materials-1}

Using the \emph{rvest} library in the statistical package \emph{R}, we
pulled body text for individual articles from each of the aforementioned
10 news sources (identified using CSS language). We compiled the
articles into a dataset (Wickham, 2016). Using this dataset, we
identified word count and average word count per source. This process
was run for articles pertaining to Kavanaugh's nomination that were
published between September 13, 2018 and October 11, 2018 inclusive.
This date range was selected in reference to the widely-publicized and
viewed nomination hearing on September 27, 2018. We set the start date
at September 13 (two weeks before the hearing) and the end date at
October 11 (two weeks after the hearing) so that we could capture a
large amount of data (roughly one month) during which Kavanaugh's
nomination was at its peak saturation in news coverage.

\subsection{Data analysis}\label{data-analysis-1}

As in Experiment 1, the text was scanned with \emph{ngram}. Again, blank
articles, text from the Disqus system, and duplicate articles were
removed (Schmidt et al., 2017). The text was processed and stemmed in
order to convert to a usable form for further analysis (Feinerer \&
Hornik, 2017). Words were subsequently matched with their valences from
Warriner et al. (2013). Depending on the results of Jordan et al.
(2019)'s research, alternative forms of the Moral Foudnations Dictionary
may be imported instead of the original dictionary.

@@

\section{Experiment 2}\label{experiment-2-1}

\subsection{Results}\label{results-1}

\section{Discussion}\label{discussion-1}

\section{Conclusions}\label{conclusions}

@@

\section{References}\label{references}

@@

\hypertarget{refs}{}
\hypertarget{ref-Aust2017}{}
Aust, F., \& Barth, M. (2017). papaja: Create APA manuscripts with R
Markdown. Retrieved from \url{https://github.com/crsh/papaja}

\hypertarget{ref-Bradley1999}{}
Bradley, M. M., \& Lang, P. J. (1999). \emph{Affective Norms for English
Words (ANEW): Instruction Manual and Affective Ratings} (No. C-1). The
Center for Research in Psychophysiology, University of Florida.

\hypertarget{ref-Cameron2006}{}
Cameron, D. (2006). Ideology and language. \emph{Journal of Political
Ideologies}, \emph{11}(2), 141--152.
doi:\href{https://doi.org/10.1080/13569310600687916}{10.1080/13569310600687916}

\hypertarget{ref-Clinton2014}{}
Clinton, J. D., \& Enamorado, T. (2014). The National News Media's
Effect on Congress: How Fox News Affected Elites in Congress. \emph{The
Journal of Politics}, \emph{76}(4), 928--943.
doi:\href{https://doi.org/10.1017/S0022381614000425}{10.1017/S0022381614000425}

\hypertarget{ref-Druckman2004}{}
Druckman, J. N., Jacobs, L. R., \& Ostermeier, E. (2004). Candidate
Strategies to Prime Issues and Image. \emph{The Journal of Politics},
\emph{66}(4), 1180--1202.
doi:\href{https://doi.org/10.1111/j.0022-3816.2004.00295.x}{10.1111/j.0022-3816.2004.00295.x}

\hypertarget{ref-Feinerer2017}{}
Feinerer, I., \& Hornik, K. (2017). Text mining package. Retrieved from
\url{http://tm.r-forge.r-project.org/}

\hypertarget{ref-Gelman2006}{}
Gelman, A. (2006). Multilevel (hierarchical) modeling: What it can and
cannot do. \emph{Technometrics}, \emph{48}(3), 432--435.
doi:\href{https://doi.org/10.1198/004017005000000661}{10.1198/004017005000000661}

\hypertarget{ref-Gilligan1982}{}
Gilligan, C. (1982). New maps of development: New visions of maturity.
\emph{American Journal of Orthopsychiatry}, \emph{52}(2), 199--212.
doi:\href{https://doi.org/10.1111/j.1939-0025.1982.tb02682.x}{10.1111/j.1939-0025.1982.tb02682.x}

\hypertarget{ref-Graham2009}{}
Graham, J., Haidt, J., \& Nosek, B. A. (2009). Liberals and
conservatives rely on different sets of moral foundations. \emph{Journal
of Personality and Social Psychology}, \emph{96}(5), 1029--1046.
doi:\href{https://doi.org/10.1037/a0015141}{10.1037/a0015141}

\hypertarget{ref-Graham2011}{}
Graham, J., Nosek, B. A., Haidt, J., Iyer, R., Koleva, S., \& Ditto, P.
H. (2011). Mapping the moral domain. \emph{Journal of Personality and
Social Psychology}, \emph{101}(2), 366--385.
doi:\href{https://doi.org/10.1037/a0021847}{10.1037/a0021847}

\hypertarget{ref-Haidt2007}{}
Haidt, J., \& Graham, J. (2007). When Morality Opposes Justice:
Conservatives Have Moral Intuitions that Liberals may not Recognize.
\emph{Social Justice Research}, \emph{20}(1), 98--116.
doi:\href{https://doi.org/10.1007/s11211-007-0034-z}{10.1007/s11211-007-0034-z}

\hypertarget{ref-Jordan2019}{}
Jordan, K. N., Buchanan, E. M., \& Padfield, W. E. (2019). \emph{A
Validation of the Moral Foundations Questionnaire and Dictionary}.
Retrieved from \url{https://osf.io/kt9yf/}

\hypertarget{ref-Kohlberg1977}{}
Kohlberg, L., \& Hersh, R. H. (1977). Moral development: A review of the
theory. \emph{Theory into Practice}, \emph{16}(2), 53--59.
doi:\href{https://doi.org/10.1080/00405847709542675}{10.1080/00405847709542675}
